---
title: "Direct exposure experiment: effects of amendments and contamination" 
author: "RTD, II"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## set working directory
setwd("../Agri_Contaminants")

## packages
library("tidyverse") ## ggplot, stringr, and more
library("knitr") ## include_graphics function
library("kableExtra") ## fancy tables
library("lme4") ## lmer function
library("car") ## Anova for type 3 (interactions)
library("emmeans") ## post-hoc analyses
library("RColorBrewer") ## colour pallets
library("cowplot") ## plotgrid function
library("ggbeeswarm") ## geom_beeswarm package
#library(broom) ## tidy() function
library("pheatmap") ## plotting correlation matrices
library("caret") ## findCorrelation func
```

## Direct impacts of amendinants on crop growth

What impacts do amendments spiked with different levels of
contaminants have on plant traits, and are these impacts crop-species dependent?

Experimental design:

* Crossed two amendment types (biosolids [BS] and reclaimed water [RW]) 
with three amendinant spiking levels (0 [no spiking], 1 [ecologically relevant 
concentrations], and 2 [higher than ecological concentrations, to detect in 
plant tissues downstream]).
* Exposed the resulting six treatments to each of three crop species 
commonly used in agriculture (lettuce, radish, and green peas), each growing 
separately in pots, resulting in a total of 18 treatment combinations 
(2 amendments x 3 spiking levels x 3 crop species). 
* To account for environmental heterogeneity in growth conditions across the 
greenhouse, each treatment combination was spatially replicated three times, 
for a total of 54 experimental units (i.e., pots). 
* Finally, we included nine controls in which no amendments were added, for a 
total of 63 pots (n = 3 controls per crop species x 3 crops = 9 pots). 
    
Traits measured:

* Post-planting: 
  + Germination rates (proportion of germinated seeds out of total number of 
      seeds planted)
  + Survival rates (proportion of plants that survived out of total number of 
      seeds planted)    
* At harvest: 
  + Normalized Difference Vegetation Index (NDVI) value, which reflects leaf 
      chlorophyll A content and is a close correlate of nitrogen content, 
  + Above- and below-ground wet and dry biomass 
* For peas only: 
  + Wet and dry pod weights (the portion of the plant most 
      relevant to agricultural yield). 
  + Number of flowers and pods 
  + Infection rate (proportion of plants infected by fungus out of total)
      
### Load data and format

```{r direct_load_data}
## load raw data
direct <- read_csv("./raw_data/direct-updated.csv")

## calculate other traits
direct <- direct %>%
  mutate(seeds_planted_total = 
           rowSums(across(c(seeds_planted1, seeds_planted2)),
                                       na.rm=TRUE),
         seeds_germinated_total = 
           rowSums(across(c(seeds_germinated,
                            seeds_germinated2)),
                                       na.rm=TRUE),
         germination_perc.r = (seeds_germinated_total / 
                               seeds_planted_total)*100,
         germination_perc = ifelse(germination_perc.r > 100,
                                   100,
                                   germination_perc.r),
         survival_perc.r = (plants / seeds_planted_total) * 100,
         survival_perc = ifelse(survival_perc.r > 100,
                                100,
                                survival_perc.r),
         dry_total_weight = rowSums(across(c(dry_shoot_weight, 
                                             dry_root_weight))),
         shoot_root_ratio = dry_shoot_weight / dry_root_weight,
         shoot_moisture = (wet_shoot_weight - dry_shoot_weight) /
           dry_shoot_weight,
         per_plant_weight = dry_total_weight / plants,
         infected_peas_perc = 
           (infected_peas/seeds_planted_total)*100
         )

### change weight pod weight to NA for samples without pods
direct$wet_pod_weight.corr <- ifelse(direct$pods > 0, 
                                     direct$wet_pod_weight,
                                     NA)

direct$dry_pod_weight.corr <- ifelse(direct$pods > 0, 
                                     direct$dry_pod_weight,
                                     NA)

direct$per_pod_weight <- ifelse(is.na(direct$pods) == FALSE,
                                direct$dry_pod_weight.corr / direct$pods,
                                NA)
  
### add 1's to survival percentage (to take log)
direct$survival_perc.corr <- direct$survival_perc + 1

### add 1's to percent infected
direct$infected_peas_perc.corr <- 
  ifelse(is.na(direct$infected_peas_perc) == 
           FALSE,
         direct$infected_peas_perc + 1,
         NA)

## add in NDVI data
ndvi <- read_csv("./raw_data/direct_biomass_ndvi.csv")

ndvi$NDVI_mean <- 
  rowMeans(ndvi[, c("ndvi_1", "ndvi_2", "ndvi_3")], 
                           na.rm = TRUE)

direct$NDVI_mean <- ndvi$NDVI_mean[match(direct$sample_ID, 
                                      ndvi$sample_ID)]
direct$NDVI_mean <- ifelse(is.nan(direct$NDVI_mean) == TRUE,
                           NA,
                           direct$NDVI_mean)

## Code factors

### specify spiking level as an ordered factor
direct$spikeFac <- factor(direct$spiking_level,
                          levels = c("0","1","2"),
                          ordered = TRUE)

### rename source to be consistent with indirect
direct$amend <- ifelse(
  is.na(direct$source) == TRUE, "control",
                        ifelse(
                          direct$source ==
                            "effluents","RW",
                               "BS"))

direct$amend <- factor(direct$amend,
                        levels = c(
                          "control",
                          "RW",
                          "BS"
                        ))

### spiking level plus amend
direct$amend_spike <- paste0(direct$amend, 
                                "-", 
                                direct$spiking_level)

### treatment
direct$amend_spike <- factor(direct$amend_spike, 
                                levels = c("control-0",
                                           "RW-0",
                                           "RW-1",
                                           "RW-2",
                                           "BS-0",
                                           "BS-1",
                                           "BS-2"
                                           ),
                              ordered = TRUE)

### species
direct$species <- factor(direct$species, 
                                levels = c("lettuce",
                                           "radish",
                                           "pea"))

## save formatted data file
save(direct, file = "./direct_files/direct-formatted.Rda")

## write csv file
write.csv(direct, "./direct_files/raw_data.formatted.csv",
          row.names = FALSE)
```

### 1) Examine trait correlations

Ensuring we are picking traits that are not highly correlated

```{r trait_corrs}
## set output directory
dir <- "./direct_files/figs" 

## load required packages
library(readr)
library(ggplot2)
library(dplyr)
library(pheatmap)
library(factoextra)
library(FactoMineR)
library(cowplot)
library(tidyr)
library(tibble)
library(purrr)

## load data
load(file = "./direct_files/direct-formatted.Rda") ## loads direct

## specify trait vars (all species)
trait.list.all <- c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "shoot_root_ratio",
                "germination_perc",
                "survival_perc",
                ## pea specific
                "wet_pod_weight",
                "dry_pod_weight",
                "per_pod_weight",
                "infected_peas_perc",
                ## "NDVI_mean", ## too many NAs due to low survival
                "pods",
                "flowers")

### select all traits
direct_cols_df <- direct[, c(trait.list.all, "species", "sample_ID")]

# ## replace NAs with 0s for all
# direct_cols_df <- direct_cols_df %>%
#   mutate(across(where(is.numeric), ~ ifelse(is.na(.), 0, .)))

## run the following function to examine correlations across and within species
source("./Source_code/1_direct-plot_corr.func.R")

all_traits <- c("wet_shoot_weight", "dry_shoot_weight", "wet_root_weight", "dry_root_weight",
            "dry_total_weight", "per_plant_weight", "shoot_moisture", "shoot_root_ratio",
            "germination_perc", "survival_perc")

pea_traits <- c("wet_shoot_weight", "dry_shoot_weight", "wet_root_weight", "dry_root_weight",
            "dry_total_weight", "per_plant_weight", "shoot_moisture", "shoot_root_ratio",
            "germination_perc", "survival_perc",
            "wet_pod_weight", "dry_pod_weight", "per_pod_weight",
            "infected_peas_perc", "pods", "flowers")

vars_by_species <- list(
  "ALL" = all_traits,
  "lettuce" = all_traits,
  "radish" = all_traits,
  "pea" = pea_traits)

result <- run_corr(data = direct_cols_df,
                   species_col = "species",
                   output_dir = dir,
                   top_n_vars = 3,
                   vars_by_species = vars_by_species)
```

### Direct exposure experiment: models

#### 2) Model: amendment x species (SP-0)

Impact of amendment type on plant traits, compared across 
different crop species

Include controls. 
Limit to Spiking level 0 (no additional contaminants added)

```{r m1-species.vs.amend}
## load data
load(file = "./direct_files/direct-formatted.Rda") ## loads direct

## set contrasts for ANOVA
options(contrasts=c("contr.sum","contr.poly")) 

### traits to run through
trait.list <- c(
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "shoot_root_ratio",
                "germination_perc",
                "survival_perc.corr",
                ### pea specific
                "dry_pod_weight.corr",
                "per_pod_weight",
                "NDVI_mean",
                "pods",
                "flowers"
                )

# Create output directory if it doesn't exist
## set output directory
output_dir <- "./direct_files/models/"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

## source function
source("./Source_code/2_direct-mod_species.vs.amend.func.R")

## model
mod.out <- sapply(trait.list, 
                   mod_species.vs.amend,
                   df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## combine dfs
### ANOVAs
aov <- lapply(mod.out, `[[`, 1) %>%
  bind_rows(.)
### format
aov$pval <- ifelse(is.na(aov$`Pr(>F)`) == FALSE,
                   signif(aov$`Pr(>F)`, 3),
                   signif(aov$`Pr(>Chisq)`, 3))
aov$sig <- ifelse(aov$pval< 0.001, ", P < 0.001",
            ifelse(aov$pval < 0.01, paste0(", P = ", aov$pval, "**"),
            ifelse(aov$pval < 0.05, paste0(", P = ", aov$pval, "*"),
            ifelse(aov$pval < 0.1, paste0(", P = ", aov$pval, "."),
                               ", P > 0.1"))))
aov$stat <- ifelse(is.na(aov$`F value`) == FALSE,
                        signif(aov$`F value`, 3),
                        signif(aov$`LR Chisq`, 3))

## pivot wider for each model term
aov.w <- aov %>%
  select(-c(`F value`,"Pr(>F)", "LR Chisq", 
            "Pr(>Chisq)","pval", "Sum Sq")) %>%
  pivot_wider(
    names_from = term,
    values_from = c(stat, Df, sig)
  ) ## 15 traits total, 15 rows

## formatting
aov.f <- aov.w %>%
  mutate(Amendment = ifelse(trait %in% c("pods","flowers"),
                         paste0("X2 = ", stat_amend, 
                                ", df = ", Df_amend,
                                sig_amend),
                         paste0("F [", Df_amend, ", ",
                                Df_Residuals, "] = ", stat_amend,
                                sig_amend)),
         Species = ifelse(trait %in% c(
                               "dry_pod_weight.corr", 
                               "per_pod_weight",
                               "NDVI_mean",
                               "pods",
                               "flowers"), 
                          "N/A (Pea only)",
                    paste0("F [", Df_species, ", ",
                           Df_Residuals, "] = ", stat_species,
                           sig_species)),
         `Species x Amendment` = 
           ifelse(trait %in% c(
                               "dry_pod_weight.corr", 
                               "per_pod_weight",
                               "NDVI_mean",
                               "pods",
                               "flowers"), 
                                  "N/A (Pea only)",
                        paste0("F [", 
                               `Df_species:amend`, ", ",
                                Df_Residuals, "] = ",
                               `stat_species:amend`,
                               `sig_species:amend`))
         ) %>%
  select(trait, Species, Amendment, `Species x Amendment`)

### CONTRASTS (amendment within species)
cont <- lapply(mod.out, `[[`, 3) %>%
  bind_rows(.)

## Amendment (regardless of species)
cont.amend.germ <- mod.out[["germination_perc"]][[4]]
cont.amend.surv <- mod.out[["survival_perc.corr"]][[4]]
cont <- rbind(cont, cont.amend.germ, cont.amend.surv)

## add in pea
cont$crop <- ifelse(is.na(cont$species) == TRUE, "pea", 
                    ifelse(cont$species == ".", "all",
                           paste0(cont$species)))
## round pval
cont$pval <- signif(cont$p.value, 3)
### format
cont$sig <- ifelse(cont$p.value < 0.001, ", P < 0.001",
            ifelse(cont$p.value < 0.01, paste0(", P = ", cont$pval, "**"),
            ifelse(cont$p.value < 0.05, paste0(", P = ", cont$pval, "*"),
            ifelse(cont$p.value < 0.1, paste0(", P = ", cont$pval, "."),
                               ", P > 0.1"))))
## test stats
cont$stat <- ifelse(is.na(cont$t.ratio) == FALSE,
                        signif(cont$t.ratio, 3),
                        signif(cont$z.ratio, 3))

cont$stat.f <- ifelse(cont$trait %in% c("pods","flowers"),
                      paste0("z = ", cont$stat,
                    cont$sig),
                    paste0("t = ", cont$stat, ", ", 
                    "df = ", cont$df,
                    cont$sig)
)
## estimates
cont$log2FC <- log2(cont$ratio)
cont$LCL <- ifelse(is.na(cont$lower.CL) == TRUE,
                   signif(log2(cont$asymp.LCL), 3),
                   signif(log2(cont$lower.CL), 3))
cont$UCL <- ifelse(is.na(cont$upper.CL) == TRUE,
                   signif(log2(cont$asymp.UCL), 3),
                   signif(log2(cont$upper.CL), 3))
cont$log2FC_CL <- paste0(signif(cont$log2FC, 3), " [",
                        cont$LCL," to ",
                        cont$UCL,"]")
## save as R data file (full version)
save(cont, file = paste0(output_dir, "cont1-species.vs.amend.Rdata"))
## save as csv (Supp file)
cont.s <- cont %>%
  mutate(amend = ifelse(grepl("RW", contrast, fixed = FALSE) == TRUE,
                        "RW", "BS")) %>%
  select(trait, crop, amend, log2FC_CL, stat.f)
write_csv(cont.s, file = paste0(output_dir, "cont1-species.vs.amend.csv"))

## save relevant output for table
cont.f <- cont %>%
  filter(p.value < 0.1) %>%
  mutate(
    contrast_info = paste0(crop, "_", contrast, ": ", 
                           stat.f, "; ", 
                           "log2FC [95% CL] = ", log2FC_CL)
    ) %>%
  select(trait, contrast_info) %>%
  group_by(trait) %>%
  summarize(
    sig_contrasts = paste(contrast_info, collapse = " | ")
    )

## add to anova
aov.f$sig_contrasts <- cont.f$sig_contrasts[match(aov.f$trait, cont.f$trait)] 
## save
write.csv(aov.f, file = paste0(output_dir, "aov1-species.vs.amend.csv"), 
          row.names = FALSE)
## print nicely
knitr::kable(
 aov.f,
  format = "markdown",   # ensures Markdown-friendly table
  digits = 3,
  align = "c",
  caption = "ANOVA (T3SS) for amendment and species impacts on performance"
)


### emmeans
emm <- lapply(mod.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, file = paste0(output_dir, "emm1-species.vs.amend.csv"), 
          row.names = FALSE)
```

#### 3) Model: species x contamination level 

Impact of contamination concentration on plant traits, 
compared within amendment types

Exclude controls (non-spiked) 

```{r m2-species.vs.contam}
## load data
load(file = "./direct_files/direct-formatted.Rda") ## loads direct

## set contrasts for ANOVA
options(contrasts=c("contr.sum","contr.poly")) 

### traits to run through
### traits to run through
trait.list <- c(
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "shoot_root_ratio",
                "germination_perc",
                "survival_perc.corr",
                ### pea specific
                "dry_pod_weight.corr",
                "per_pod_weight",
                ## "NDVI_mean", ## excluded due to not being measured in 2/3 BS treats
                "pods",
                "flowers"
                )

## create list of traits/amends to loop through:
amend.list <- c("RW","BS")
comb_df <- expand_grid(traits = trait.list,
                       amends = amend.list)
## Combine into a single string
comb_df <- comb_df %>%
  mutate(
    combs = paste(traits, amends, sep = "_")
  )
### 28 combinations: 14 traits x 2 amends = 28

# Create output directory if it doesn't exist
## set output directory
output_dir <- "./direct_files/models/"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

## source the function:
source("./Source_code/3_direct-mod_species.vs.contam.func.R")

mod.out <- mapply(FUN = mod_species.vs.contam, 
                      combs = comb_df$combs, 
                      traits = comb_df$traits, 
                      amends = comb_df$amends, 
                      USE.NAMES = TRUE, 
                      SIMPLIFY = FALSE,
                      MoreArgs = list(df = direct))

## combine dfs
### ANOVAs
aov <- lapply(mod.out, `[[`, 1) %>%
  bind_rows(.)

### format
aov$pval <- ifelse(is.na(aov$`Pr(>F)`) == FALSE,
                   signif(aov$`Pr(>F)`, 3),
                   signif(aov$`Pr(>Chisq)`, 3))
aov$sig <- ifelse(aov$pval< 0.001, ", P < 0.001",
            ifelse(aov$pval < 0.01, paste0(", P = ", aov$pval, "**"),
            ifelse(aov$pval < 0.05, paste0(", P = ", aov$pval, "*"),
            ifelse(aov$pval < 0.1, paste0(", P = ", aov$pval, "."),
                               ", P > 0.1"))))
aov$stat <- ifelse(is.na(aov$`F value`) == FALSE,
                        signif(aov$`F value`, 3),
                        signif(aov$`LR Chisq`, 3))

## pivot wider for each model term
aov.w <- aov %>%
  select(-c(`F value`,"Pr(>F)", "LR Chisq", 
            "Pr(>Chisq)","pval", "Sum Sq")) %>%
  pivot_wider(
    names_from = term,
    values_from = c(stat, Df, sig)
  ) ## 14 traits by two amend types = 28 rows

## formatting
aov.f <- aov.w %>%
  mutate(`Spiking level (SL)` = ifelse(trait %in% c("pods","flowers"),
                         paste0("X2 = ", stat_spikeFac, 
                                ", df = ", Df_spikeFac,
                                sig_spikeFac),
                         paste0("F [", Df_spikeFac, ", ",
                                Df_Residuals, "] = ", stat_spikeFac,
                                sig_spikeFac)),
         Species = ifelse(trait %in% c("wet_pod_weight.corr",
                               "dry_pod_weight.corr", 
                               "per_pod_weight",
                               "NDVI_mean",
                               "pods",
                               "flowers"), "N/A (Pea only)",
                    paste0("F [", Df_species, ", ",
                           Df_Residuals, "] = ", stat_species,
                           sig_species)),
         `Species x SL` = 
           ifelse(trait %in% c("wet_pod_weight.corr",
                               "dry_pod_weight.corr", 
                               "NDVI_mean",
                               "per_pod_weight",
                               "pods",
                               "flowers"), "N/A (Pea only)",
                        paste0("F [", 
                               `Df_species:spikeFac`, ", ",
                                Df_Residuals, "] = ",
                               `stat_species:spikeFac`,
                               `sig_species:spikeFac`))
         ) %>%
  select(trait, Amendment = amend, Species, 
         `Spiking level (SL)`, 
         `Species x SL`)

### CONTRASTS
cont <- lapply(mod.out, `[[`, 3) %>%
  bind_rows(.)

## Spiking level (regardless of species)
cont.SL.germ.RW <- mod.out[["germination_perc_RW"]][[4]]
cont.SL.germ.BS <- mod.out[["germination_perc_BS"]][[4]]
cont.SL.surv.RW <- mod.out[["survival_perc.corr_RW"]][[4]]
cont.SL.surv.BS <- mod.out[["survival_perc.corr_BS"]][[4]]

## combine
cont <- rbind(cont, cont.SL.germ.RW, cont.SL.surv.RW,
              cont.SL.germ.BS, cont.SL.surv.BS)

## add in pea
cont$crop <- ifelse(is.na(cont$species) == TRUE, "pea", 
                    ifelse(cont$species == ".", "all",
                           paste0(cont$species)))
## round pval
cont$pval <- signif(cont$p.value, 3)
### format
cont$sig <- ifelse(cont$p.value < 0.001, ", P < 0.001",
            ifelse(cont$p.value < 0.01, paste0(", P = ", cont$pval, "**"),
            ifelse(cont$p.value < 0.05, paste0(", P = ", cont$pval, "*"),
            ifelse(cont$p.value < 0.1, paste0(", P = ", cont$pval, "."),
                               ", P > 0.1"))))
## test stats
cont$stat <- ifelse(is.na(cont$t.ratio) == FALSE,
                        signif(cont$t.ratio, 3),
                        signif(cont$z.ratio, 3))

cont$stat.f <- ifelse(cont$trait %in% c("pods","flowers"),
                      paste0("z = ", cont$stat,
                    cont$sig),
                    paste0("t = ", cont$stat, ", ", 
                    "df = ", cont$df,
                    cont$sig)
)
## estimates
cont$Est_r <- signif(cont$estimate, 3)
cont$LCL <- ifelse(is.na(cont$lower.CL) == TRUE,
                   signif(cont$asymp.LCL, 3),
                   signif(cont$lower.CL, 3))
cont$UCL <- ifelse(is.na(cont$upper.CL) == TRUE,
                   signif(cont$asymp.UCL, 3),
                   signif(cont$upper.CL, 3))
cont$Est_CL <- paste0(cont$Est_r, " [",
                        cont$LCL," to ",
                        cont$UCL,"]")
## save as R data file
save(cont, file = paste0(output_dir, "cont2-species.vs.contam.Rdata"))
## save as csv
cont.w <- cont %>%
  select(contrast, amend, crop, trait, Est_CL, stat.f) %>%
  pivot_wider(
    names_from = "contrast",
    values_from = c("Est_CL", "stat.f")
  ) %>%
  select(trait, crop, amend, Est_CL_linear, stat.f_linear,
          Est_CL_quadratic, stat.f_quadratic)

cont1 <- read_csv(file = paste0(output_dir, "cont1-species.vs.amend.csv"))

cont.s <- cont1 %>%
  left_join(., cont.w, by = c("trait", "crop", "amend"))
  
write_csv(cont.s, file = paste0("Supplementary_files/File_S1-direct.csv"))

## save relevant output for table
cont.f <- cont %>%
  filter(p.value < 0.1) %>%
  mutate(
    contrast_info = paste0(crop, "_", contrast, ": ", 
                           stat.f, "; ", 
                           "Est. [95% CL] = ", Est_CL)
    ) %>%
  select(trait, Amendment = amend, contrast_info) %>%
  group_by(trait, Amendment) %>%
  summarize(
    sig_contrasts = paste(contrast_info, collapse = " | ")
    )

## add to anova
aov.f <- left_join(aov.f, cont.f, by = c("trait","Amendment"))
## save
write.csv(aov.f, file = paste0(output_dir, "aov2-species.vs.contam.csv"), 
          row.names = FALSE)
## print nicely
knitr::kable(
  aov.f,
  format = "markdown",   # ensures Markdown-friendly table
  digits = 3,
  align = "c",
  caption = "ANOVA (T3SS) for contaminant and species impacts on performance"
)

### emmeans (back-transformed manually)
emm <- lapply(mod.out, `[[`, 2) %>%
  bind_rows(.)
emm$emmean.bt <- exp(emm$emmean)
emm$LCL.bt <- ifelse(is.na(emm$lower.CL) == TRUE,
                         exp(emm$asymp.LCL),
                         exp(emm$lower.CL))
emm$UCL.bt <- ifelse(is.na(emm$upper.CL) == TRUE,
                         exp(emm$asymp.UCL),
                         exp(emm$upper.CL))
## save
write.csv(emm, file = paste0(output_dir, "emm2-species.vs.contam.csv"), 
          row.names = FALSE)
```

#### 4) Models: infection rate and survival

* M3: Is infection rate impacted by contaminant level?
* M4: Is survival rate impacted by infection rate?

```{r direct_m3_4}
# Create output directory if it doesn't exist
## set output directory
output_dir <- "./direct_files/models/"
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

## load data
load(file = "./direct_files/direct-formatted.Rda") ## loads direct

## set contrasts for ANOVA
options(contrasts=c("contr.sum","contr.poly")) 

## infection rate and contaminant levels
m3 <- lm(log(infected_peas_perc.corr) ~ spikeFac, 
         data = direct %>%
           filter(species == "pea" & amend == "BS"))
m3_sum <- summary(m3)
## res vs fits plots
png(paste0(output_dir, "resfits_plots/m3-infect.vs.contam.png"), 
      width=6, height=6, units='in', res=300)
  layout(matrix(1:4, ncol = 2))
  plot(m3)
  layout(1)
dev.off()
## format output
aov <- as.data.frame(Anova(m3, type = 2))
aov$term <- rownames(aov)
aov$trait <- "infection_rate"
aov$amend <- "BS"
### format
aov$pval <- signif(aov$`Pr(>F)`, 3)
aov$sig <- ifelse(aov$pval< 0.001, ", P < 0.001",
            ifelse(aov$pval < 0.01, paste0(", P = ", aov$pval, "**"),
            ifelse(aov$pval < 0.05, paste0(", P = ", aov$pval, "*"),
            ifelse(aov$pval < 0.1, paste0(", P = ", aov$pval, "."),
                               ", p > 0.1"))))
aov$stat <- signif(aov$`F value`, 3)

## pivot wider for each model term
aov.w <- aov %>%
  select(-c(`F value`,"Pr(>F)", "pval", "Sum Sq")) %>%
  pivot_wider(
    names_from = term,
    values_from = c(stat, Df, sig)
  ) ## 1 traits by one amend types = 1 row

## formatting
aov.f <- aov.w %>%
  mutate(`Spiking level (SL)` = paste0("F [", Df_spikeFac, ", ",
                                Df_Residuals, "] = ",
                                stat_spikeFac,
                                sig_spikeFac),
         ) %>%
  select(trait, Amendment = amend, `Spiking level (SL)`)

### CONTRASTS
emm <- emmeans(m3, poly ~ spikeFac,
                       infer = TRUE)
cont <- as.data.frame(emm$contrasts)
cont$trait <- "infection_rate"
cont$amend <- "BS"
## round pval
cont$pval <- signif(cont$p.value, 3)
### format
cont$sig <- ifelse(cont$p.value < 0.001, ", P < 0.001",
            ifelse(cont$p.value < 0.01, paste0(", P = ", cont$pval, "**"),
            ifelse(cont$p.value < 0.05, paste0(", P = ", cont$pval, "*"),
            ifelse(cont$p.value < 0.1, paste0(", P = ", cont$pval, "."),
                               ", p > 0.1"))))
## test stats
cont$stat <- signif(cont$t.ratio, 3)

cont$stat.f <- paste0("t = ", cont$stat, ", ", 
                    "df = ", cont$df,
                    cont$sig)
## estimates
cont$Est_r <- signif(cont$estimate, 3)
cont$LCL <- signif(cont$lower.CL, 3)
cont$UCL <- signif(cont$upper.CL, 3)
cont$Est_CL <- paste0(cont$Est_r, " [",
                        cont$LCL," to ",
                        cont$UCL,"]")
## save relevant output for table
cont.f <- cont %>%
  filter(p.value < 0.1 &
           contrast == "linear") %>%
  mutate(
    contrast_info = paste0(contrast, ": ", 
                           stat.f, "; ", 
                           "Est. [95% CL] = ", Est_CL)
    ) %>%
  select(trait, Amendment = amend, contrast_info) %>%
  group_by(trait, Amendment) %>%
  summarize(
    sig_contrasts = paste(contrast_info, collapse = " | ")
    )

## add to anova
aov.f <- left_join(aov.f, cont.f, by = c("trait","Amendment"))
## save
write.csv(aov.f, file = paste0(output_dir, "m3-infect.vs.contam.csv"), 
          row.names = FALSE)
kable(aov.f)

## infection rate and survival
## set contrasts (for linear regression)
options(contrasts=c("contr.treatment","contr.poly")) 
## model
m4 <- lm(survival_perc ~ infected_peas_perc, 
         data = direct %>%
           filter(species == "pea" & amend == "BS"))
m4_sum <- summary(m4)
png(paste0(output_dir, "resfits_plots/m4-surv.vs.infect.png"), 
      width=6, height=6, units='in', res=300)
  layout(matrix(1:4, ncol = 2))
  plot(m4)
  layout(1)
dev.off()
aov <- Anova(m4, type = 2)
aov$term <- rownames(aov)
aov$trait <- "survival_rate"
aov$amend <- "BS"
### format
aov$pval <- signif(aov$`Pr(>F)`, 3)
aov$sig <- ifelse(aov$pval< 0.001, ", P < 0.001",
            ifelse(aov$pval < 0.01, paste0(", P = ", aov$pval, "**"),
            ifelse(aov$pval < 0.05, paste0(", P = ", aov$pval, "*"),
            ifelse(aov$pval < 0.1, paste0(", P = ", aov$pval, "."),
                               ", p > 0.1"))))
aov$stat <- signif(aov$`F value`, 3)

## pivot wider for each model term
aov.w <- aov %>%
  select(-c(`F value`,"Pr(>F)", "pval", "Sum Sq")) %>%
  pivot_wider(
    names_from = term,
    values_from = c(stat, Df, sig)
  ) ## 1 traits by one amend types = 1 row

## formatting
aov.f <- aov.w %>%
  mutate(`Infection (%)` = paste0("F [", 
                                  Df_infected_peas_perc, ", ",
                                Df_Residuals, "] = ",
                                stat_infected_peas_perc,
                                sig_infected_peas_perc),
         ) %>%
  select(trait, Amendment = amend, `Infection (%)`)
## get model coeffs
m4_coeff <- as.data.frame(m4_sum[["coefficients"]])
m4_coeff$term <- rownames(m4_coeff)
m4_coeff$trait <- "survival_rate"
m4_coeff <- m4_coeff %>%
  filter(term != "(Intercept)")
m4_coeff$`linear reg.` <- paste0("t = ",
  signif(m4_coeff$`t value`, 3), ", P = ",
  signif(m4_coeff$`Pr(>|t|)`, 3), "; Slope +/- SE = ",
  signif(m4_coeff$Estimate, 3), " +/- ",
  signif(m4_coeff$`Std. Error`, 3))
## merge together
m4_aov <- left_join(aov.f, 
                    m4_coeff[,c("trait","linear reg.")], 
                    by = "trait")
## save
write.csv(m4_aov, paste0(output_dir, "m4-surv.vs.infect.csv"), 
          row.names = TRUE)
kable(m4_aov)
```

### Figures (Direct)

#### 5) Fig: raw data

```{r plot_raw_data}
### raw data
load(file = "./direct_files/direct-formatted.Rda") ## loads direct

## source function
source("./Source_code/4_direct-plot_raw_data.func.R")

## plots (traits at harvest)

### traits to run through
trait.list <- c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "shoot_root_ratio",
                "germination_perc",
                "survival_perc",
                ### pea specific
                "wet_pod_weight",
                "dry_pod_weight",
                "per_pod_weight",
                "infected_peas_perc",
                "NDVI_mean",
                "pods",
                "flowers"
                )

## plotting function
plots.out <- sapply(trait.list, 
                     plot_raw_data,
                     df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

##### Raw data (all crops) #####
fig1 <- plot_grid(plots.out[["dry_total_weight"]] +
                    labs(y = "Dry total \n biomass (g)"),
                  plots.out[["per_plant_weight"]] +
                    labs(y = "Per plant \n biomass (g)"),
                   plots.out[["shoot_root_ratio"]] +
                    labs(y = "Shoot:root ratio"),
                  plots.out[["shoot_moisture"]] +
                    labs(y = "Shoot moisture \n (%)"),
                  plots.out[["germination_perc"]] +
                    labs(y = "Germination \n (%)"),
          ncol = 1,
          nrow = 5,
          align = "v",
          labels = NULL)
fig1

## extract just the legend (for final fig)
legend <- get_legend(
  plots.out[["dry_pod_weight"]] +
    theme(legend.position = "bottom"))

## add a outline around plot
fig1.box <- ggdraw() +
  draw_plot(fig1, x = 0.02, y = 0.02, width = 0.96, height = 0.96) +
  theme_void() +
  theme(plot.background = element_rect(color = "black", 
                                       linewidth = 2, 
                                       fill = NA))
fig1.box

##### Raw data (peas only) #####

##### Infection by contaminant level (peas only) #####

## model 3: infection and contaminant level
direct.f <- direct %>%
         filter(species == "pea" & amend == "BS") %>%
  droplevels(.)

### means
df.means <- direct.f %>% 
  group_by(spikeFac) %>%
  summarize(mean = mean(infected_peas_perc, 
                        na.rm = TRUE))

## plot
M3_plot <- ggplot(df.means,
                  aes(x = spikeFac, y = mean)) +
    geom_bar(aes(fill = spikeFac),
             stat = "identity") +
    geom_beeswarm(data = direct.f,
                  aes(y = infected_peas_perc)) +
    scale_fill_manual(values = c("#EABD8C",
                                 "#FFAD00",
                                 "#B06500")) +
    labs(x = "Contamination level", 
         y = "Infected (%)") +
    guides(fill = "none") +
    theme_bw() +
    theme(
      axis.text.x = element_text(size = 12),
      axis.title.x = element_text(size = 14),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 14, face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
    )
M3_plot

## model 4: infection and survival
##### Infection vs Survival (peas only) #####
M4_plot <- ggplot(direct %>%
         filter(species == "pea" & amend == "BS"),
       aes(x = infected_peas_perc, y = survival_perc)) +
  geom_smooth(method = "lm", colour = "black") +
  geom_point(aes(colour = spikeFac), size = 3) +
  scale_colour_manual(values = c("#EABD8C",
                                 "#FFAD00",
                                 "#B06500")) +
  labs(y = "Survival (%)", 
         x = "Infected (%)") +
    guides(colour = "none") +
    theme_bw() +
    theme(
      axis.text.x = element_text(size = 12),
      axis.title.x = element_text(size = 14),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 14, face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
    )
M4_plot

## plot grid for just the pea traits
peas_top <- plot_grid(
  plots.out[["dry_pod_weight"]] + ## TL 
    labs(y = "Dry pod weight (g)"),
   plots.out[["per_pod_weight"]] + ## TL 
    labs(y = "Per pod weight (g)"),
 plots.out[["pods"]] + ## TR
    labs(y = "Pods (no.)"),
  plots.out[["flowers"]] + ## ML
    labs(y = "Flowers (no.)"), 
    M3_plot,
    M4_plot,
  ncol = 2,
  nrow = 3,
  align = "hv",
  axis = "tblr",
  labels = NULL)
peas_top

## put the legend along the bottom
legend_bottom <- plot_grid(legend, ncol = 1)
legend_bottom

## put them together
fig_peas <- plot_grid(
  peas_top,
  legend_bottom,
  nrow = 2,
  ncol = 1,
  rel_heights = c(1, 0.1),
  align = "v"
)
fig_peas

## add an outline around fig
fig_peas.box <- ggdraw() +
  draw_plot(fig_peas, x = 0.02, y = 0.02, width = 0.96, height = 0.96) +
  theme_void() +
  theme(plot.background = element_rect(color = "black", 
                                       linewidth = 2, 
                                       fill = NA))
fig_peas.box

## plot grid for all traits
fig_all <- plot_grid(
          fig1.box, 
          fig_peas.box,
          ncol = 2,
          rel_widths = c(1, 1),
          align = "hv",
          axis = "tblr",
          labels = c("A", "B"))
fig_all

## add titles for each column (to specify species)

# Create a row of subtitles

# Create a row of four subtitles
x_coords <- seq(1/12, 11/12, length.out = 6)
subtitle_row <- ggdraw() +
  draw_label("Lettuce", x = x_coords[[1]] + 0.04, y = 0.5, 
             hjust = 0.5, size = 14, fontface = "bold") +
  draw_label("Radish", x = x_coords[[2]] + 0.02, y = 0.5, 
             hjust = 0.5, size = 14, fontface = "bold") +
  draw_label("Pea", x = x_coords[[3]], y = 0.5, 
             hjust = 0.5, size = 14, fontface = "bold") +
  draw_label("Pea", x = x_coords[[5]], y = 0.5, 
             hjust = 0.5, size = 14, fontface = "bold")

# Combine subtitle row and main panel
final_plot <- plot_grid(
  subtitle_row,
  fig_all,
  ncol = 1,
  rel_heights = c(0.02, 1)
)
final_plot

## save
ggsave("./direct_files/figs/Fig1-direct.png", 
       width = 10, height = 8, 
       units = "in")
```

#### 6) Post-hoc tests (supplementary figs)

```{r figs_PHT}
## figures (contrasts)
load(file = "./direct_files/models/cont1-species.vs.amend.Rdata") ## loads cont

## formatting
cont$amend <- ifelse(grepl("BS", cont$contrast, fixed = FALSE),
                      "BS", "RW")

cont$amend <- factor(cont$amend,
                      levels = c("RW",
                                 "BS"))

### sig traits
sig.traits <- c("dry_total_weight",
                "per_plant_weight",
                "shoot_root_ratio",
                "shoot_moisture",
                #"survival_perc.corr"
                "germination_perc"
                )

### filter dataset to sig traits
cont.f <- cont %>%
         filter(trait %in% sig.traits) %>%
         droplevels(.)

cont.f$trait <- factor(cont.f$trait,
          levels = c(#"dry_shoot_weight","dry_root_weight",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_root_ratio",
                "shoot_moisture",
                #"survival_perc.corr"
                "germination_perc"
                ))

### rename traits
trait_names <- c(
  per_plant_weight = 'Per plant \n biomass (g)',
  dry_shoot_weight = 'Dry shoot \n biomass (g)',
  dry_total_weight = 'Dry total \n biomass (g)',
  dry_root_weight = 'Dry root \n biomass (g)',
  shoot_root_ratio = 'Shoot to \n root ratio',
  shoot_moisture = 'Shoot \n moisture \n (%)',
  germination_perc = 'Germination \n (%)',
  survival_perc.corr = 'Survival \n (%)',
  pods = "Pods (no.)"
)

species_names <- c(
  lettuce = 'Lettuce',
  pea = 'Pea',
  radish = 'Radish'
)

## drop "." for species
cont.fd <- cont.f %>%
  filter(species %in% c("lettuce", "pea", "radish")) %>%
  droplevels(.)

##### Contrasts (amendment type) #####
M1_plot <- ggplot(data = cont.fd,
   aes(x = amend, 
       y = log2(ratio), 
       colour = amend)) +
  geom_pointrange(aes(ymin = log2(lower.CL),
                      ymax = log2(upper.CL)),
                  position = position_dodge(0.5)) +
  geom_hline(aes(yintercept = 0),
             linetype = 2) +
  ### add in significance (0.05 <= p < 0.1) 
  geom_text(
    data = cont.fd %>% filter(pval >= 0.05 & pval < 0.1),
    aes(x = amend, y = log2FC),
    label = "+",
    size = 6,
    position = position_dodge(0.5),
    vjust = -0.5
  ) +
  ### add in significance (p < 0.05) 
  geom_text(
    data = cont.fd %>% filter(pval < 0.05),
    aes(x = amend, y = log2FC),
    label = "*",
    size = 6,
    position = position_dodge(0.5),
    vjust = -0.3
  ) +
  coord_flip() +
   scale_colour_manual(values = c(
                                  "blue",
                                 "#B06500")) +
  facet_grid(trait~species, scales = "fixed",
             labeller = labeller(trait = trait_names,
                                 species = species_names)) +
  labs(x = "Amendment", y = expression(log[2]~"Fold Change (95% CL)")) +
  guides(colour = "none") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title.x = element_text(size =  14),
    axis.title.y = element_text(size =  14),
    strip.text.y = element_blank(),
    strip.text.x = element_text(size = 12, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )
M1_plot

## M2

## emms (emmeans)
emms <- read_csv("./direct_files/models/emm2-species.vs.contam.csv")

## linear contrasts (to get sig linear trends)
load(file = "./direct_files/models/cont2-species.vs.contam.Rdata")

## include non-linear & linear terms
cont.w <- cont %>%
  pivot_wider(
    names_from = contrast,
    values_from = c("Est_CL", "pval"),
    id_cols =  c("crop","amend","trait")
  )

## combine emms and contrasts
emms <- emms %>%
  left_join(
    cont.w %>% select(species = crop, everything()),
    by = c("species","amend","trait")
  )

## format
emms$amend <- factor(emms$amend,
                      levels = c("RW",
                                 "BS"))

emms$species <- ifelse(is.na(emms$species) == TRUE,
                       "pea", emms$species)
emms$species <- factor(emms$species,
                       levels = c("lettuce",
                                  "radish",
                                  "pea"))

### filter dataset to sig traits
emms.f <- emms %>%
         filter(trait %in% sig.traits) %>%
         droplevels(.)

## trait order
emms.f$trait <- factor(emms.f$trait,
          levels = c(
         "dry_total_weight",
                "per_plant_weight",
                "shoot_root_ratio",
                "shoot_moisture",
                #"survival_perc.corr"
         "germination_perc"
         ))

##### Contrasts (contamination level) #####
M2_plot <- ggplot(data = emms.f,
   aes(x = factor(spikeFac), 
       y = emmean, 
       colour = amend)) +
  
  geom_line(aes(group = amend),
            position = position_dodge(0.2)) +
  
  geom_pointrange(data = emms.f %>% filter(trait != "pods"),
                  aes(ymin = lower.CL, ymax = upper.CL),
                  position = position_dodge(0.2)) +
  
  geom_pointrange(data = emms.f %>% filter(trait == "pods"),
                  aes(ymin = asymp.LCL, ymax = asymp.UCL),
                  position = position_dodge(0.2)) +
  
  ### linear: marginal significance
  geom_text(
    data = emms.f %>%
      filter(spikeFac == 2 & pval_linear >= 0.05 & pval_linear < 0.1),
    aes(x = factor(spikeFac), y = emmean),
    label = "+", size = 6,
    position = position_dodge(0.5),
    vjust = -0.5
  ) +
  
  ### linear: significant
  geom_text(
    data = emms.f %>%
      filter(spikeFac == 2 & pval_linear < 0.05),
    aes(x = factor(spikeFac), y = emmean),
    label = "*", size = 6,
    position = position_dodge(0.5),
    vjust = -0.5
  ) +
  
  ### quadratic: significant
  geom_text(
    data = emms.f %>%
      filter(spikeFac == 0 & pval_quadratic < 0.05),
    aes(x = factor(spikeFac), y = emmean),
    label = "*", size = 6,
    position = position_dodge(0.5),
    vjust = -0.5
  ) +
  
  ### quadratic: marginal significance
  geom_text(
    data = emms.f %>%
      filter(spikeFac == 0 & pval_quadratic >= 0.05 & pval_quadratic < 0.1),
    aes(x = factor(spikeFac), y = emmean),
    label = "+", size = 6,
    position = position_dodge(0.5),
    vjust = -0.5
  ) +

  scale_colour_manual(values = c("blue", "#B06500")) +

  facet_grid(trait ~ species, scales = "free",
             labeller = labeller(trait = trait_names,
                                 species = species_names)) +

  labs(y = expression("Estimated Mean (log scale)"), 
       x = "Contaminant level") +
  
  guides(colour = "none") +

  ### key parts to ensure annotation visibility
  scale_y_continuous(expand = c(0.05, 0.25)) + 
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(t = 12, r = 8, b = 8, l = 8)) +

  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    strip.text = element_text(size = 12, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

M2_plot

plots <- plot_grid(M1_plot, M2_plot,
                   ncol = 2,
                   rel_widths = c(0.8, 1),
                   align = "h",
                   labels = c("A","B"))
plots
## save
ggsave("./direct_files/figs/plots-PHT.png", 
       width = 10, height = 7, 
       units = "in")
```

