---
title: "stat_analyses"
author: "RTD, II"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## set working directory
setwd("../Agri_Contaminants")

## packages
library(tidyverse) ## ggplot, stringr, and more
library(lme4) ## lmer function
library(car) ## Anova for type 3 (interactions)
library(emmeans) ## post-hoc analyses
library(RColorBrewer) ## colour pallets
library(cowplot) ## plotgrid function
library(ggbeeswarm) ## geom_beeswarm package
```

## Direct impacts of contaminants on crop growth

What impacts do amendments spiked with different levels of
contaminants have on plant traits, and are these impacts crop-species dependent?

Experimental design:

* Crossed two amendment types (biosolids [BS] and reclaimed water [RW]) with 
three contaminant spiking levels (0 [no spiking], 1 [ecologically relevant
concentrations], and 2 [higher than ecological concentrations, to detect in 
plant tissues downstream]).
* Exposed the resulting six treatments to each of three crop species 
commonly used in agriculture (lettuce, radish, and green peas), each growing 
separately in pots, resulting in a total of 18 treatment combinations 
(2 amendments x 3 spiking levels x 3 crop species). 
* To account for environmental heterogeneity in growth conditions across the 
greenhouse, each treatment combination was spatially replicated three times, 
for a total of 54 experimental units (i.e., pots). 
* Finally, we included nine controls in which no amendments were added, for a 
total of 63 pots (n = 3 controls per crop species x 3 crops = 9 pots). 
    
Traits measured:

* Post-planting: 
  + Germination rates (proportion of germinated seeds out of total number of 
      seeds planted)
* At harvest: 
  + Normalized Difference Vegetation Index (NDVI) value, which reflects leaf 
      chlorophyll A content and is a close correlate of nitrogen content, 
  + Above- and below-ground wet and dry biomass 
  + Wet and dry pod weights for peas only (the portion of the plant most 
      relevant to agricultural yield). 
      
### Load data

```{r load_data}
direct <- read_csv("./direct-updated.csv")

## calculate other traits
direct <- direct %>%
  mutate(seeds_planted_total = 
           rowSums(across(c(seeds_planted1, seeds_planted2)),
                                       na.rm=TRUE),
         seeds_germinated_total = 
           rowSums(across(c(seeds_germinated, seeds_germinated2)),
                                       na.rm=TRUE),
         germination_perc = (seeds_germinated_total / 
                               seeds_planted_total)*100,
         dry_total_weight = rowSums(across(c(dry_shoot_weight, 
                                             dry_root_weight))),
         shoot_root_ratio = dry_shoot_weight / dry_root_weight,
         shoot_moisture = (wet_shoot_weight - dry_shoot_weight) / 
           dry_shoot_weight,
         per_plant_weight = dry_total_weight / plants,
         survival_perc = (plants / seeds_planted_total) * 100,
         infected_peas_perc = (infected_peas/seeds_planted_total)*100
         )

## Code factors

### specify spiking level as an ordered factor
direct$spikeFac <- factor(direct$spiking_level,
                          levels = c("0","1","2"),
                          ordered = TRUE)

### rename source to be consistent with indirect
direct$contam <- ifelse(
  is.na(direct$source) == TRUE, "control",
                        ifelse(
                          direct$source ==
                            "effluents","RW",
                               "BS"))

### spiking level plus contam
direct$contam_spike <- paste0(direct$contam, 
                                "-", 
                                direct$spiking_level)

### treatment
direct$contam_spike <- factor(direct$contam_spike, 
                                levels = c("control-0",
                                           "BS-0",
                                           "BS-1",
                                           "BS-2",
                                           "RW-0",
                                           "RW-1",
                                           "RW-2"))

## save formatted data file
save(direct, file = "./direct-formatted.Rda")
```

### Figures (raw data)

```{r figs_raw}
load(file = "./direct-formatted.Rda") ## loads direct

## plots (traits at harvest)

### traits to run through (all crops)
trait.list1 <- c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "shoot_root_ratio",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "germination_perc",
                "survival_perc"
                )

## plotting function
plots.out1 <- sapply(trait.list1, df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE,
                  function(traits, df){
                    
### print traits
  print(traits)

### means
  df.means <- df %>% 
    group_by(species, contam_spike, contam, spikeFac) %>%
    summarize(mean = mean(get(traits), na.rm = TRUE))
  controls <- df %>% 
    filter(contam == "control") %>%
    group_by(species, contam_spike, contam, spikeFac) %>%
    summarize(mean = mean(get(traits), na.rm = TRUE))

### plots
  p <- ggplot(df.means,
       aes(x = contam_spike, 
           y = mean)) +
  geom_bar(aes(fill = contam_spike),
           stat = "identity") +
  geom_beeswarm(data = df,
             aes(y = get(traits))
             ) +
   geom_hline(data = controls,
              aes(yintercept = mean),
              linetype = 2) +
  facet_wrap(~species, scales = "free_y") +
  scale_fill_manual(values = c("gray",
                                "#EABD8C",
                                  "#FFAD00",
                                  "#B06500",
                                 "lightblue",
                                 "cornflowerblue",
                                 "blue")) +
  labs(y = paste0(traits), 
       x = "Contaminant source") +
  guides(fill = "none") +
  theme_bw() +
  theme(
    axis.text.x = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_text(size = 16),
    strip.text = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )
  
### return
  return(p)
})

## peas only

### traits to run through (peas only)
trait.list2 <- c(
                "wet_pod_weight",
                "dry_pod_weight",
                "pods",
                "flowers",
                "infected_peas_perc"
                )

## plotting function
plots.out2 <- sapply(trait.list2, df = direct %>%
                       filter(species == "pea") %>%
                       droplevels(.),
                   simplify = FALSE, 
                   USE.NAMES = TRUE,
                  function(traits, df){
                    
### print traits
  print(traits)

### means
  df.means <- df %>% 
    group_by(species, contam_spike, contam, spikeFac) %>%
    summarize(mean = mean(get(traits), na.rm = TRUE))
  controls <- df %>% 
    filter(contam == "control") %>%
    group_by(species, contam_spike, contam, spikeFac) %>%
    summarize(mean = mean(get(traits), na.rm = TRUE))

### plots
  p <- ggplot(df.means,
       aes(x = contam_spike, 
           y = mean)) +
  geom_bar(aes(fill = contam_spike),
           stat = "identity") +
  geom_beeswarm(data = df,
             aes(y = get(traits))
             ) +
   geom_hline(data = controls,
              aes(yintercept = mean),
              linetype = 2) +
  # facet_wrap(~species, scales = "free_y") +
   scale_fill_manual(values = c("gray",
                                "#EABD8C",
                                  "#FFAD00",
                                  "#B06500",
                                 "lightblue",
                                 "cornflowerblue",
                                 "blue")) +
  labs(y = paste0(traits), 
       x = "Treatment") +
  guides(fill = "none") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )
  
### return
  return(p)
})

## combine plots (all crops)
fig1 <- plot_grid(plots.out1[["dry_shoot_weight"]] +
                    labs(y = "Dry shoot \n weight (g)"),
                  plots.out1[["dry_root_weight"]] +
                    labs(y = "Dry root \n weight (g)"),
                  plots.out1[["shoot_root_ratio"]] +
                    labs(y = "Shoot:root \n ratio"),
                  plots.out1[["shoot_moisture"]] +
                    labs(y = "Shoot moisture"),
                  plots.out1[["survival_perc"]] +
                    labs(y = "Survival (%)"),
          ncol = 1,
          nrow = 5,
          align = "hv",
          labels = NULL)

## combine plots (peas only)
fig2 <- plot_grid(plots.out2[["dry_pod_weight"]] +
                    labs(y = "Dry pod \n weight (g)"),
                  plots.out2[["infected_peas_perc"]] +
                    labs(y = "Peas Infected (%)"),
          ncol = 2,
          nrow = 1,
          align = "hv",
          labels = NULL)

## combine plots
fig <- plot_grid(fig1, fig2,
          ncol = 1,
          nrow = 2,
          rel_heights = c(0.8,0.2),
          align = "hv",
          labels = NULL)

save_plot("./raw_data-direct.png", fig,
          ncol = 2, 
          nrow = 5,
          base_aspect_ratio = 1.3)
```

## Models

### Model 1

Impact of amendment and contaminant exposure on plant traits, compared across 
different crop species

Include controls. Interaction with species only

NOTE: consensus from data lunch (04April2025)

Because spiking levels are very different between the amendments, we cannot 
include the interaction terms.

```{r model1}
## load data
load(file = "./direct-formatted.Rda") ## loads direct

## set contrasts for ANOVA
options(contrasts=c("contr.sum","contr.poly")) 

### models (test)
lmm <- lm(log(dry_shoot_weight) ~ species*contam_spike, 
            data = direct)

### traits to run through (non-peas)
trait.list1 <- c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "shoot_root_ratio",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "germination_perc",
                "survival_perc"
                )


## model
mod1.out <- sapply(trait.list1, 
                   df = direct %>%
                     filter(survival_perc > 0), ## cannot take a log of 0
                   simplify = FALSE, 
                   USE.NAMES = TRUE,
                  function(traits, df){
                    
### print traits
  print(traits)
### model
  lmm <- lm(log(get(traits)) ~ contam_spike*species,
              data = df)
### res vs fit plot
  png(paste0("./model_outputs/direct/resfits1_", 
             traits, ".png"), 
      width=6, height=6, units='in', res=300)
  layout(matrix(1:4, ncol = 2))
  plot(lmm)
  layout(1)
  dev.off()
### model output
  print(summary(lmm))
  aov <- Anova(lmm, type = 3)
### processing
  aov$trait <- paste0(traits)
  aov$term <- rownames(aov)

### compare treatments within species
  lmm.bt <- update(ref_grid(lmm), tran = "log")
  lmm.emm <- emmeans(lmm.bt, trt.vs.ctrl ~ 
                       contam_spike | species,
                     infer = TRUE,
                     type = "response")

### processing
  emm <- as.data.frame(lmm.emm$emmeans)
  emm$trait <- paste0(traits)
  cont <- as.data.frame(lmm.emm$contrasts)
  cont$trait <- paste0(traits)
  
### return dfs
  return(list(aov, ## [[1]]
              emm, ## [[2]]
              cont ## [[3]]
              )) 
                    
})

## combine dfs
### ANOVAs
aov <- lapply(mod1.out, `[[`, 1) %>%
  bind_rows(.)
write.csv(aov, "./model_outputs/direct/aov1.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod1.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/direct/emm1.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod1.out, `[[`, 3) %>%
  bind_rows(.)
write.csv(cont, "./model_outputs/direct/cont1.csv", 
          row.names = FALSE)

## figures
cont <- read_csv("model_outputs/direct/cont1.csv")

## rename treatment
cont$treat <- ifelse(grepl("BS-0", 
                           cont$contrast, 
                           fixed = FALSE), "BS-0", 
                    ifelse(grepl("BS-1", 
                                 cont$contrast, 
                                 fixed = FALSE), "BS-1",
                    ifelse(grepl("BS-2", 
                                 cont$contrast, 
                                 fixed = FALSE), "BS-2",
                    ifelse(grepl("RW-0", 
                                 cont$contrast, 
                                 fixed = FALSE), "RW-0",
                    ifelse(grepl("RW-1", 
                                 cont$contrast, 
                                 fixed = FALSE), "RW-1",
                           "RW-2")))))

### all traits

## EMmeans

## figures
emms <- read_csv("model_outputs/direct/emm1.csv")

### extract spiking level
emms$spike <- sub(".*-", "", emms$contam_spike)
emms$amend <- sub("-.*", "", emms$contam_spike)

### plot
ggplot(emms %>% filter(trait %in% c(
  "dry_shoot_weight","dry_root_weight",
  "shoot_root_ratio","shoot_moisture",
  "survival_perc") &
    amend != "control"),
   aes(x = spike, 
       y = response, 
       colour = amend)) +
  geom_pointrange(aes(ymin = lower.CL,
                      ymax = upper.CL),
                  position = position_dodge(0.5)) +
   geom_hline(data = emms %>%
                filter(contam_spike == "control-0"),
                       aes(yintercept = response),
              linetype = 2) +
   scale_colour_manual(values = c(
                                  "#B06500",
                                 "blue")) +
  facet_wrap(~trait, scales = "free_y", ncol = 1) +
  labs(y = "Ratio", x = "Species") +
  guides(colour = "none") +
  theme_bw() +
  theme(
    axis.text.x = element_text(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )


### plot
ggplot(cont %>% filter(trait %in% c(
  "dry_shoot_weight","dry_root_weight",
  "shoot_root_ratio","shoot_moisture",
  "survival_perc"
    )),
   aes(x = species, 
       y = ratio, 
       colour = contrast)) +
  geom_pointrange(aes(ymin = lower.CL,
                      ymax = upper.CL),
                  position = position_dodge(0.5)) +
   geom_hline(aes(yintercept = 1),
              linetype = 2) +
   scale_colour_manual(values = c(
                                "#EABD8C",
                                  "#FFAD00",
                                  "#B06500",
                                 "lightblue",
                                 "cornflowerblue",
                                 "blue")) +
  facet_wrap(~trait, scales = "free_y", ncol = 1) +
  labs(y = "Ratio", x = "Species") +
  guides(colour = "none") +
  theme_bw() +
  theme(
    axis.text.x = element_text(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

ggsave("./model_outputs/direct/cont1.png", width = 12, height = 8, units = "in")

## peas only

### traits to run through (peas only)
trait.list2 <- c("wet_pod_weight",
                "dry_pod_weight",
                "pods",
                "flowers",
                "infected_peas_perc"
                )

## model
mod2.out <- sapply(trait.list2, df = direct %>%
                     filter(species == "pea"),
                   simplify = FALSE, 
                   USE.NAMES = TRUE,
                  function(traits, df){
                    
### print traits
  print(traits)
### model
  lmm <- lm(get(traits) ~ contam_spike,
              data = df)
### res vs fit plot
  png(paste0("./model_outputs/direct/resfits2_", 
             traits, ".png"), 
      width=6, height=6, units='in', res=300)
  layout(matrix(1:4, ncol = 2))
  plot(lmm)
  layout(1)
  dev.off()
### model output
  print(summary(lmm))
  aov <- Anova(lmm, type = 2)
### processing
  aov$trait <- paste0(traits)
  aov$term <- rownames(aov)

### compare treatments within species
  #lmm.bt <- update(ref_grid(lmm), tran = "log")
  lmm.emm <- emmeans(lmm, trt.vs.ctrl ~ 
                       contam_spike,
                     infer = TRUE)

### processing
  emm <- as.data.frame(lmm.emm$emmeans)
  emm$trait <- paste0(traits)
  cont <- as.data.frame(lmm.emm$contrasts)
  cont$trait <- paste0(traits)
  
### return dfs
  return(list(aov, ## [[1]]
              emm, ## [[2]]
              cont ## [[3]]
              )) 
                    
})

## combine dfs
### ANOVAs
aov <- lapply(mod2.out, `[[`, 1) %>%
  bind_rows(.)
write.csv(aov, "./model_outputs/direct/aov2.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod2.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/direct/emm2.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod2.out, `[[`, 3) %>%
  bind_rows(.)
write.csv(cont, "./model_outputs/direct/cont2.csv", 
          row.names = FALSE)

## figures
cont <- read_csv("model_outputs/direct/cont2.csv")

## rename treatment
cont$treat <- ifelse(grepl("BS-0", 
                           cont$contrast, 
                           fixed = FALSE), "BS-0", 
                    ifelse(grepl("BS-1", 
                                 cont$contrast, 
                                 fixed = FALSE), "BS-1",
                    ifelse(grepl("BS-2", 
                                 cont$contrast, 
                                 fixed = FALSE), "BS-2",
                    ifelse(grepl("RW-0", 
                                 cont$contrast, 
                                 fixed = FALSE), "RW-0",
                    ifelse(grepl("RW-1", 
                                 cont$contrast, 
                                 fixed = FALSE), "RW-1",
                           "RW-2")))))

### pea traits

### plot
ggplot(cont,
       aes(x = treat, 
           y = estimate, 
           colour = contrast)) +
  geom_pointrange(aes(ymin = lower.CL,
                      ymax = upper.CL),
                  position = position_dodge(0.5)) +
   geom_hline(aes(yintercept = 0),
              linetype = 2) +
   scale_colour_manual(values = c(
                                "#EABD8C",
                                  "#FFAD00",
                                  "#B06500",
                                 "lightblue",
                                 "cornflowerblue",
                                 "blue")) +
  facet_wrap(~trait, scales = "free_y", ncol = 2) +
  labs(y = "Estimate", x = "Treatment") +
  guides(colour = "none") +
  theme_bw() +
  theme(
    axis.text.x = element_text(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

ggsave("./model_outputs/direct/cont1.png", width = 12, height = 8, units = "in")



```
