---
title: "stat_analyses"
author: "RTD, II"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## set working directory
setwd("../Agri_Contaminants")

## packages
library(tidyverse) ## ggplot, stringr, and more
library(lme4) ## lmer function
library(car) ## Anova for type 3 (interactions)
library(emmeans) ## post-hoc analyses
library(RColorBrewer) ## colour pallets
library(cowplot) ## plotgrid function
library(ggbeeswarm) ## geom_beeswarm package
```

## Direct impacts of contaminants on crop growth

What impacts do amendments spiked with different levels of
contaminants have on plant traits, and are these impacts crop-species dependent?

Experimental design:

* Crossed two amendment types (biosolids [BS] and reclaimed water [RW]) 
with three contaminant spiking levels (0 [no spiking], 1 [ecologically relevant 
concentrations], and 2 [higher than ecological concentrations, to detect in 
plant tissues downstream]).
* Exposed the resulting six treatments to each of three crop species 
commonly used in agriculture (lettuce, radish, and green peas), each growing 
separately in pots, resulting in a total of 18 treatment combinations 
(2 amendments x 3 spiking levels x 3 crop species). 
* To account for environmental heterogeneity in growth conditions across the 
greenhouse, each treatment combination was spatially replicated three times, 
for a total of 54 experimental units (i.e., pots). 
* Finally, we included nine controls in which no amendments were added, for a 
total of 63 pots (n = 3 controls per crop species x 3 crops = 9 pots). 
    
Traits measured:

* Post-planting: 
  + Germination rates (proportion of germinated seeds out of total number of 
      seeds planted)
  + Survival rates (proportion of plants that survived out of total number of 
      seeds planted)    
* At harvest: 
  + Normalized Difference Vegetation Index (NDVI) value, which reflects leaf 
      chlorophyll A content and is a close correlate of nitrogen content, 
  + Above- and below-ground wet and dry biomass 
* For peas only: 
  + Wet and dry pod weights (the portion of the plant most 
      relevant to agricultural yield). 
  + Number of flowers and pods 
  + Infection rate (proportion of plants infected by fungus out of total)
      
### Load data

```{r direct_load_data}
direct <- read_csv("./direct-updated.csv")

## calculate other traits
direct <- direct %>%
  mutate(seeds_planted_total = 
           rowSums(across(c(seeds_planted1, seeds_planted2)),
                                       na.rm=TRUE),
         seeds_germinated_total = 
           rowSums(across(c(seeds_germinated, seeds_germinated2)),
                                       na.rm=TRUE),
         germination_perc = (seeds_germinated_total / 
                               seeds_planted_total)*100,
         dry_total_weight = rowSums(across(c(dry_shoot_weight, 
                                             dry_root_weight))),
         shoot_root_ratio = dry_shoot_weight / dry_root_weight,
         shoot_moisture = (wet_shoot_weight - dry_shoot_weight) / 
           dry_shoot_weight,
         per_plant_weight = dry_total_weight / plants,
         survival_perc = (plants / seeds_planted_total) * 100,
         infected_peas_perc = (infected_peas/seeds_planted_total)*100
         )

### change weight pod weight to NA for samples without pods
direct$wet_pod_weight.corr <- ifelse(direct$pods > 0, 
                                     direct$wet_pod_weight,
                                     NA)

direct$dry_pod_weight.corr <- ifelse(direct$pods > 0, 
                                     direct$dry_pod_weight,
                                     NA)

### add 1's to survival (to take log)
direct$survival_perc.corr <- direct$survival_perc + 1

### add 1's to percent infected
direct$infected_peas_perc.corr <- 
  ifelse(is.na(direct$infected_peas_perc) == 
           FALSE,
         direct$infected_peas_perc + 1,
         NA)

## Code factors

### specify spiking level as an ordered factor
direct$spikeFac <- factor(direct$spiking_level,
                          levels = c("0","1","2"),
                          ordered = TRUE)

### rename source to be consistent with indirect
direct$contam <- ifelse(
  is.na(direct$source) == TRUE, "control",
                        ifelse(
                          direct$source ==
                            "effluents","RW",
                               "BS"))

direct$contam <- factor(direct$contam,
                        levels = c(
                          "control",
                          "RW",
                          "BS"
                        ))

### spiking level plus contam
direct$contam_spike <- paste0(direct$contam, 
                                "-", 
                                direct$spiking_level)

### treatment
direct$contam_spike <- factor(direct$contam_spike, 
                                levels = c("control-0",
                                           "RW-0",
                                           "RW-1",
                                           "RW-2",
                                           "BS-0",
                                           "BS-1",
                                           "BS-2"
                                           ),
                              ordered = TRUE)

## save formatted data file
save(direct, file = "./direct-formatted.Rda")
```

### Models

#### Model 1: amendment type x species

Impact of amendment type on plant traits, compared across 
different crop species

Include controls. 
Limit to Spiking level 0 (no additional contaminants added)

```{r direct_m1}
## load data
load(file = "./direct-formatted.Rda") ## loads direct

## set contrasts for ANOVA
options(contrasts=c("contr.sum","contr.poly")) 

### traits to run through (non-peas)
trait.list <- c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "shoot_root_ratio",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "germination_perc",
                "survival_perc.corr",
                ### pea specific
                "wet_pod_weight.corr",
                "dry_pod_weight.corr",
                "pods",
                "flowers"
                )

## source function
source("./Source_code/lmm1_func.R")

## model
mod1.out <- sapply(trait.list, 
                   lmm1_func,
                   df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## combine dfs
### ANOVAs
aov <- lapply(mod1.out, `[[`, 1) %>%
  bind_rows(.)
### format
aov$pval <- ifelse(is.na(aov$`Pr(>F)`) == FALSE,
                   aov$`Pr(>F)`,
                   aov$`Pr(>Chisq)`)
aov$sig <- ifelse(aov$pval< 0.001, "***",
            ifelse(aov$pval < 0.01, "**",
            ifelse(aov$pval < 0.05, "*",
            ifelse(aov$pval < 0.1, ".",
                               "ns"))))
aov$stat <- ifelse(is.na(aov$`F value`) == FALSE,
                   signif(aov$`F value`, 3),
                   signif(aov$`LR Chisq`, 3))
aov$stat_sig <- paste0(aov$stat, aov$sig)
write.csv(aov, "./model_outputs/direct/aov1.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod1.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/direct/emm1.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod1.out, `[[`, 3) %>%
  bind_rows(.)
cont$crop <- ifelse(is.na(cont$species) == TRUE, "pea", 
                    paste0(cont$species))
### format
cont$sig <- ifelse(cont$'p.value' 
                     < 0.001, "***",
                    ifelse(cont$'p.value' 
                           < 0.01, "**",
                    ifelse(cont$'p.value' 
                                  < 0.05, "*",
                    ifelse(cont$'p.value' 
                                       < 0.1,
                                       ".",
                                       "ns"))))
cont$LCL <- ifelse(is.na(cont$lower.CL) == TRUE,
                   signif(cont$asymp.LCL, 3),
                   signif(cont$lower.CL, 3))
cont$UCL <- ifelse(is.na(cont$upper.CL) == TRUE,
                   signif(cont$asymp.UCL, 3),
                   signif(cont$upper.CL, 3))
cont$ratio_CL <- paste0(signif(cont$ratio, 3)," (",
                        cont$LCL," - ",
                        cont$UCL,")")
write.csv(cont, "./model_outputs/direct/cont1.csv", 
          row.names = FALSE)
```

#### Model 2: contamination level x species

Impact of contamination level on plant traits, compared within amendment types

Exclude controls (non-amended) 

```{r direct_m2}
## load data
load(file = "./direct-formatted.Rda") ## loads direct

## set contrasts for ANOVA
options(contrasts=c("contr.sum","contr.poly")) 

### traits to run through
trait.list2 <- rep(c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "shoot_root_ratio",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "germination_perc",
                "survival_perc.corr",
                ### pea specific
                "wet_pod_weight.corr",
                "dry_pod_weight.corr",
                "pods",
                "flowers"
                ), 2)

## create list of traits/amends to loop through:
amend.list <- c(rep(c("RW"),14), rep(c("BS"),14))
comb.list <- paste0(trait.list2, ", ", amend.list)

## source the function:
source("./Source_code/lmm2_func.R")

mod2.out <- mapply(FUN = lmm2_func, 
                      combs = comb.list, 
                      traits = trait.list2, 
                      amends = amend.list, 
                      USE.NAMES = TRUE, 
                      SIMPLIFY = FALSE,
                      MoreArgs = list(df = direct))

## combine dfs
### ANOVAs
aov <- lapply(mod2.out, `[[`, 1) %>%
  bind_rows(.)
### format
aov$pval <- ifelse(is.na(aov$`Pr(>F)`) == FALSE,
                   aov$`Pr(>F)`,
                   aov$`Pr(>Chisq)`)
aov$sig <- ifelse(aov$pval< 0.001, "***",
            ifelse(aov$pval < 0.01, "**",
            ifelse(aov$pval < 0.05, "*",
            ifelse(aov$pval < 0.1, ".",
                               "ns"))))
aov$stat <- ifelse(is.na(aov$`F value`) == FALSE,
                   signif(aov$`F value`, 3),
                   signif(aov$`LR Chisq`, 3))
aov$stat_sig <- paste0(aov$stat, aov$sig)
write.csv(aov, "./model_outputs/direct/aov2.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod2.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/direct/emm2.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod2.out, `[[`, 3) %>%
  bind_rows(.)
cont$crop <- ifelse(is.na(cont$species) == TRUE, "pea", 
                    paste0(cont$species))
### format
cont$sig <- ifelse(cont$'p.value' 
                     < 0.001, "***",
                    ifelse(cont$'p.value' 
                           < 0.01, "**",
                    ifelse(cont$'p.value' 
                                  < 0.05, "*",
                    ifelse(cont$'p.value' 
                                       < 0.1,
                                       ".",
                                       "ns"))))
cont$LCL <- ifelse(is.na(cont$lower.CL) == TRUE,
                   signif(cont$asymp.LCL, 3),
                   signif(cont$lower.CL, 3))
cont$UCL <- ifelse(is.na(cont$upper.CL) == TRUE,
                   signif(cont$asymp.UCL, 3),
                   signif(cont$upper.CL, 3))
cont$est_CL <- paste0(signif(cont$estimate, 3)," (",
                        cont$LCL," to ",
                        cont$UCL,")")
write.csv(cont, "./model_outputs/direct/cont2.csv", 
          row.names = FALSE)
```

#### Models 3 & 4: infection rate

* M3: Is infection rate impacted by contaminant level?
* M4: Is survival rate impacted by infection rate?

```{r direct_m3_4}
## set contrasts
options(contrasts=c("contr.treatment","contr.poly")) 

## load data
load(file = "./direct-formatted.Rda") ## loads direct

## infection rate and contaminant levels
m3 <- lm(log(infected_peas_perc.corr) ~ spikeFac, 
         data = direct %>%
           filter(species == "pea" & contam == "BS"))
m3_sum <- summary(m3)
png(paste0("./model_outputs/direct/resfits_m3.png"), 
      width=6, height=6, units='in', res=300)
  layout(matrix(1:4, ncol = 2))
  plot(m3)
  layout(1)
dev.off()
plot(m3)
m3_coeff <- as.data.frame(m3_sum[["coefficients"]])
write.csv(m3_coeff, "./model_outputs/direct/m3_coeff.csv", 
          row.names = TRUE)
## infection rate and survival
m4 <- lm(survival_perc ~ infected_peas_perc, 
         data = direct %>%
           filter(species == "pea" & contam == "BS"))
m4_sum <- summary(m4)
png(paste0("./model_outputs/direct/resfits_m4.png"), 
      width=6, height=6, units='in', res=300)
  layout(matrix(1:4, ncol = 2))
  plot(m4)
  layout(1)
dev.off()
m4_coeff <- as.data.frame(m4_sum[["coefficients"]])
write.csv(m4_coeff, "./model_outputs/direct/m4_coeff.csv", 
          row.names = TRUE)
```

### Figures

Raw data, emmeans, contrasts 

```{r direct_figs}
### raw data
load(file = "./direct-formatted.Rda") ## loads direct

## plots (traits at harvest)

### traits to run through (all crops)
trait.list1 <- c(
            ### all species
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "shoot_root_ratio",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "germination_perc",
                "survival_perc",
            ### pea-specific
                "wet_pod_weight",
                "dry_pod_weight",
                "pods",
                "flowers",
                "infected_peas_perc"
                )

## source function
source("./Source_code/plot_raw_data_func.R")

## plotting function
plots.out1 <- sapply(trait.list1, 
                     plot_raw_data_func,
                     df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## combine plots (all crops)
fig1 <- plot_grid(plots.out1[["dry_total_weight"]] +
                    labs(y = "Total dry weight (g)"),
                  plots.out1[["shoot_root_ratio"]] +
                    labs(y = "Shoot to root ratio"),
                  plots.out1[["shoot_moisture"]] +
                    labs(y = "Shoot moisture (%)"),
                  plots.out1[["survival_perc"]] +
                    labs(y = "Survival (%)"),
          ncol = 1,
          nrow = 4,
          align = "v",
          labels = NULL)
fig1

## combine plots (peas only)
fig2 <- plot_grid(plots.out1[["dry_pod_weight"]] +
                    labs(y = "Dry pod weight (g)"),
                  plots.out1[["pods"]] +
                    labs(y = "Pods (no.)"),
                   plots.out1[["flowers"]] +
                    labs(y = "Flowers (no.)"),
                  plots.out1[["infected_peas_perc"]] +
                    labs(y = "Peas Infected (%)"),
          ncol = 2,
          nrow = 2,
          align = "hv",
          labels = NULL)
fig2

### M1
## figures (contrasts)
cont <- read_csv("model_outputs/direct/cont1.csv")

cont$contam <- ifelse(grepl("BS", cont$contrast, fixed = FALSE),
                      "BS", "RW")

cont$contam <- factor(cont$contam,
                      levels = c("RW",
                                 "BS"))

cont$species <- ifelse(is.na(cont$species) == TRUE,
                       "pea",cont$species)

### sig traits
sig.traits <- c("survival_perc.corr",
                "shoot_moisture","shoot_root_ratio",
                "dry_total_weight")

### filter dataset to sig traits
cont.f <- cont %>%
         filter(trait %in% sig.traits) %>%
         droplevels(.)


cont.f$trait <- factor(cont.f$trait,
          levels = c(#"dry_shoot_weight","dry_root_weight",
         "dry_total_weight", "shoot_root_ratio","shoot_moisture",
         "survival_perc.corr"))

### rename traits
trait_names <- c(
  per_plant_weight = 'Per plant weight (g)',
  dry_shoot_weight = 'Dry shoot \n weight (g)',
  dry_total_weight = 'Dry total \n weight (g)',
  dry_root_weight = 'Dry root \n weight (g)',
  shoot_root_ratio = 'Shoot to root \n ratio',
  shoot_moisture = 'Shoot moisture \n (%)',
  germination_perc = 'Germination \n (%)',
  survival_perc.corr = 'Survival \n (%)',
  pods = "Pods (no.)"
)

species_names <- c(
  lettuce = 'Lettuce',
  pea = 'Pea',
  radish = 'Radish'
)

### plot
M1_plot <- ggplot(data = cont.f,
   aes(x = contam, 
       y = ratio, 
       colour = contam)) +
  geom_pointrange(aes(ymin = lower.CL,
                      ymax = upper.CL),
                  position = position_dodge(0.5)) +
  geom_hline(aes(yintercept = 1),
             linetype = 2) +
  coord_flip() +
   scale_colour_manual(values = c(
                                  "blue",
                                 "#B06500")) +
  facet_grid(trait~species, scales = "free",
             labeller = labeller(trait = trait_names,
                                 species = species_names)) +
  labs(y = NULL, x = NULL) +
  guides(colour = "none") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )
M1_plot

ggsave("./model_outputs/direct/M1_plot.png", 
       width = 8, height = 10, 
       units = "in")

## M2

## emms (emmeans)
emms <- read_csv("model_outputs/direct/emm2.csv")

emms$amend <- factor(emms$amend,
                      levels = c("RW",
                                 "BS"))

emms$species <- ifelse(is.na(emms$species) == TRUE,
                       "pea", emms$species)

### sig traits
sig.traits <- c("survival_perc.corr",
                "shoot_moisture","shoot_root_ratio",
                "dry_total_weight")

### filter dataset to sig traits
emms.f <- emms %>%
         filter(trait %in% sig.traits) %>%
         droplevels(.)

### rename traits
trait_names <- c(
  per_plant_weight = 'Per plant weight (g)',
  dry_shoot_weight = 'Dry shoot \n weight (g)',
  dry_total_weight = 'Dry total \n weight (g)',
  dry_root_weight = 'Dry root \n weight (g)',
  shoot_root_ratio = 'Shoot to root \n ratio',
  shoot_moisture = 'Shoot moisture \n (%)',
  germination_perc = 'Germination \n (%)',
  survival_perc.corr = 'Survival \n (%)',
  pods = "Pods (no.)"
)

### rename species
species_names <- c(
  lettuce = 'Lettuce',
  pea = 'Pea',
  radish = 'Radish'
)

### plot
M2_plot <- ggplot(data = emms.f,
   aes(x = spikeFac, 
       y = emmean, 
       colour = amend)) +
  geom_line(aes(group = amend),
            position = position_dodge(0.5)) +
  geom_pointrange(data = emms.f %>%
                    filter(trait != "pods"),
                           aes(ymin = lower.CL,
                      ymax = upper.CL),
                  position = position_dodge(0.5)) +
  geom_pointrange(data = emms.f %>%
                    filter(trait == "pods"),
                           aes(ymin = asymp.LCL,
                      ymax = asymp.UCL),
                  position = position_dodge(0.5)) +
   scale_colour_manual(values = c(
                                  "blue",
                                 "#B06500")) +
  facet_grid(trait~species, scales = "free",
             labeller = labeller(trait = trait_names,
                                 species = species_names)) +
  labs(y = NULL, x = NULL) +
  guides(colour = "none") +
  theme_bw() +
  theme(
    axis.text = element_text(size = 12),
    strip.text = element_text(size = 14, face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )
M2_plot

ggsave("./model_outputs/direct/M2_plot.png", 
       width = 8, height = 10, 
       units = "in")

## model 3: infection and contaminant level
### plots
direct.f <- direct %>%
         filter(species == "pea" & contam == "BS") %>%
  droplevels(.)

### means
df.means <- direct.f %>% 
  group_by(spikeFac) %>%
  summarize(mean = mean(infected_peas_perc, 
                        na.rm = TRUE))
### plot
M3_plot <- ggplot(df.means,
                  aes(x = spikeFac, y = mean)) +
    geom_bar(aes(fill = spikeFac),
             stat = "identity") +
    geom_beeswarm(data = direct.f,
                  aes(y = infected_peas_perc)) +
    scale_fill_manual(values = c("#EABD8C",
                                 "#FFAD00",
                                 "#B06500")) +
    labs(x = "Contaminant level", 
         y = "Pea Infection (%)") +
    guides(colour = "none") +
    theme_bw() +
    theme(
      axis.text.x = element_text(size = 12),
      axis.title.x = element_text(size = 16),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 16),
      strip.text = element_text(size = 14, face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
    )
M3_plot

ggsave("./model_outputs/direct/M3_plot.png", 
       width = 6, height = 6, 
       units = "in")

## model 4: infection and survival
### plots
M4_plot <- ggplot(direct %>%
         filter(species == "pea" & contam == "BS"),
       aes(x = infected_peas_perc, y = survival_perc)) +
  geom_smooth(method = "lm", colour = "black") +
  geom_point(aes(colour = spikeFac), size = 3) +
  scale_colour_manual(values = c("#EABD8C",
                                 "#FFAD00",
                                 "#B06500")) +
  labs(y = "Pea Survival (%)", 
         x = "Pea Infection (%)") +
    guides(colour = "none") +
    theme_bw() +
    theme(
      axis.text.x = element_text(size = 12),
      axis.title.x = element_text(size = 16),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 16),
      strip.text = element_text(size = 14, face = "bold"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
    )
M4_plot

ggsave("./model_outputs/direct/M4_plot.png", 
       width = 6, height = 6, 
       units = "in")
```

## Microbially-mediated impacts of contaminants

DESCRIPTION

### Load data

```{r indirect_load_data}
indirect <- read_csv("./indirect.csv")

## calculate other traits
indirect$dry_total_weight <- indirect$dry_shoot_weight + 
  indirect$dry_root_weight
indirect$shoot_root_ratio <- indirect$dry_shoot_weight / 
  indirect$dry_root_weight
indirect$shoot_moisture <- (indirect$wet_shoot_weight - 
                              indirect$dry_shoot_weight) / 
  indirect$dry_shoot_weight

## set factors

### relevel treatment
indirect$treatment <- as.factor(indirect$treatment)
indirect$treatment <- relevel(indirect$treatment, ref = "Saline Control")

### specify spiking level as an ordered factor
indirect$spikeFac <- factor(indirect$spiking_level,
                          levels = c("0","1","2"),
                          ordered = TRUE)

### rename source to be consistent with direct
indirect$contam <- factor(indirect$contam,
                        levels = c("control",
                                   "no_treat",
                                   "saline",
                          "RW",
                          "BS"
                        ))

### spiking level plus contam
indirect$contam_spike <- paste0(indirect$contam, 
                                "-", 
                                indirect$spiking_level)

### order factor
indirect$contam_spike <- factor(indirect$contam_spike, 
                                levels = c("control-0",
                                           "saline-0",
                                           "no_treat-0",
                                           "BS-0",
                                           "BS-1",
                                           "BS-2",
                                           "RW-0",
                                           "RW-1",
                                           "RW-2"))


## save formatted data file
save(indirect, file = "./indirect-formatted.Rda")

```

### Models

#### Model 1: calculate emmeans for each trait and soil slurry

Impact of prior amendment type on pea traits.
Exclude contaminants.
Include controls (from prior experiment).
No interactions.

NOTE: recommended by data lunch to calculate the geometric means from the soil
slurries, then input those in the model

```{r indirect_m1}
## loaf formatted data file
load(file = "./indirect-formatted.Rda") ## loads indirect

## set contrasts
options(contrasts=c("contr.sum","contr.poly")) 

## traits to run through
trait.list <- c("plant_height4",
                "chlorophyll_content4_mean", 
                "leaf_number2",
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "dry_total_weight",
                "shoot_root_ratio",
                "shoot_moisture")

## source function
source("./Source_code/emms_func.R")

mod1.out <- sapply(trait.list, 
                   emms_func,
                   df = indirect,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

### emmeans
emm <- lapply(mod1.out, `[[`, 1) %>%
  bind_rows(.)
save(emm, file = "./model_outputs/indirect/emm1.Rda")
### CONTRASTS
cont <- lapply(mod1.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(cont, "./model_outputs/indirect/cont1.csv", 
          row.names = FALSE)
```

#### M2: impact of prior amendment type on traits

```{r indirect_m2}
## load emmeans
load(file = "./model_outputs/indirect/emm1.Rda") ## loads emm

## set contrasts
options(contrasts=c("contr.sum","contr.poly")) 

## traits to run through
trait.list <- c("plant_height4",
                "chlorophyll_content4_mean", 
                "leaf_number2",
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "dry_total_weight",
                "shoot_root_ratio",
                "shoot_moisture")

## source function
source("./Source_code/indirect_m2.func.R")

mod2.out <- sapply(trait.list, 
                   indirect_m2.func,
                   df = emm %>%
                         filter(!contam %in% c("saline","no_treat") &
                                  spikeFac == "0") %>%
                         droplevels(.),
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## combine dfs
### ANOVAs
aov <- lapply(mod2.out, `[[`, 1) %>%
  bind_rows(.)
write.csv(aov, "./model_outputs/indirect/aov2.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod2.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/indirect/emm2.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod2.out, `[[`, 3) %>%
  bind_rows(.)
write.csv(cont, "./model_outputs/indirect/cont2.csv", 
          row.names = FALSE)
```

#### M3: impact of contaminant level

```{r indirect_m3}
## load emmeans
load(file = "./model_outputs/indirect/emm1.Rda") ## loads emm

## set contrasts
options(contrasts=c("contr.sum","contr.poly")) 

### subset data to amendment types
  df.f <- emm %>%
    filter(trait == "plant_height4" &
      contam == "RW") %>%
    droplevels(.)
  
  ### model
  lmm <- lm(log(response) ~ spikeFac,
            data = df.f)

## traits to run through
trait.list2 <- rep(c("plant_height4",
                "chlorophyll_content4_mean", 
                "leaf_number2",
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "dry_total_weight",
                "shoot_root_ratio",
                "shoot_moisture"),2)

## create list of traits/amends to loop through:
amend.list <- c(rep(c("RW"),10), rep(c("BS"),10))
comb.list <- paste0(trait.list2, ", ", amend.list)

## source function
source("./Source_code/indirect_m3.func.R")

mod3.out <- mapply(FUN = indirect_m3.func, 
                      combs = comb.list, 
                      traits = trait.list2, 
                      amends = amend.list, 
                      USE.NAMES = TRUE, 
                      SIMPLIFY = FALSE,
                      MoreArgs = list(df = emm)
                   )

## combine dfs
### ANOVAs
aov <- lapply(mod3.out, `[[`, 1) %>%
  bind_rows(.)
write.csv(aov, "./model_outputs/indirect/aov3.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod3.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/indirect/emm3.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod3.out, `[[`, 3) %>%
  bind_rows(.)
write.csv(cont, "./model_outputs/indirect/cont3.csv", 
          row.names = FALSE)
```

### Figures

```{r indirect_figs}
## loaf formatted data file
load(file = "./indirect-formatted.Rda") ## loads indirect
## load emmeans
load(file = "./model_outputs/indirect/emm1.Rda") ## loads emm

## traits to run through
trait.list <- c("plant_height4",
                "chlorophyll_content4_mean", 
                "leaf_number2",
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "dry_total_weight",
                "shoot_root_ratio",
                "shoot_moisture")

## source function
source("./Source_code/plot_emms.func.R")

plot2.out <- sapply(trait.list, 
                   plot_emms.func,
                   df = indirect %>%
                     filter(!contam %in% c("saline","no_treat")) %>%
                     droplevels(.),
                   simplify = FALSE, 
                   USE.NAMES = TRUE)
```