---
title: "stat_analyses"
author: "RTD, II"
date: "`r Sys.Date()`"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## set working directory
setwd("../Agri_Contaminants")

## packages
library(tidyverse) ## ggplot, stringr, and more
library(lme4) ## lmer function
library(car) ## Anova for type 3 (interactions)
library(emmeans) ## post-hoc analyses
library(RColorBrewer) ## colour pallets
library(cowplot) ## plotgrid function
library(ggbeeswarm) ## geom_beeswarm package
```

## Direct impacts of contaminants on crop growth

What impacts do amendments spiked with different levels of
contaminants have on plant traits, and are these impacts crop-species dependent?

Experimental design:

* Crossed two amendment types (biosolids [BS] and reclaimed water [RW]) with 
three contaminant spiking levels (0 [no spiking], 1 [ecologically relevant
concentrations], and 2 [higher than ecological concentrations, to detect in 
plant tissues downstream]).
* Exposed the resulting six treatments to each of three crop species 
commonly used in agriculture (lettuce, radish, and green peas), each growing 
separately in pots, resulting in a total of 18 treatment combinations 
(2 amendments x 3 spiking levels x 3 crop species). 
* To account for environmental heterogeneity in growth conditions across the 
greenhouse, each treatment combination was spatially replicated three times, 
for a total of 54 experimental units (i.e., pots). 
* Finally, we included nine controls in which no amendments were added, for a 
total of 63 pots (n = 3 controls per crop species x 3 crops = 9 pots). 
    
Traits measured:

* Post-planting: 
  + Germination rates (proportion of germinated seeds out of total number of 
      seeds planted)
  + Survival rates (proportion of plants that survived out of total number of 
      seeds planted)    
* At harvest: 
  + Normalized Difference Vegetation Index (NDVI) value, which reflects leaf 
      chlorophyll A content and is a close correlate of nitrogen content, 
  + Above- and below-ground wet and dry biomass 
* For peas only: 
  + Wet and dry pod weights (the portion of the plant most 
      relevant to agricultural yield). 
  + Number of flowers and pods 
  + Infection rate (proportion of plants infected by fungus out of total)
      
### Load data

```{r load_data}
direct <- read_csv("./direct-updated.csv")

## calculate other traits
direct <- direct %>%
  mutate(seeds_planted_total = 
           rowSums(across(c(seeds_planted1, seeds_planted2)),
                                       na.rm=TRUE),
         seeds_germinated_total = 
           rowSums(across(c(seeds_germinated, seeds_germinated2)),
                                       na.rm=TRUE),
         germination_perc = (seeds_germinated_total / 
                               seeds_planted_total)*100,
         dry_total_weight = rowSums(across(c(dry_shoot_weight, 
                                             dry_root_weight))),
         shoot_root_ratio = dry_shoot_weight / dry_root_weight,
         shoot_moisture = (wet_shoot_weight - dry_shoot_weight) / 
           dry_shoot_weight,
         per_plant_weight = dry_total_weight / plants,
         survival_perc = (plants / seeds_planted_total) * 100,
         infected_peas_perc = (infected_peas/seeds_planted_total)*100
         )

## Code factors

### specify spiking level as an ordered factor
direct$spikeFac <- factor(direct$spiking_level,
                          levels = c("0","1","2"),
                          ordered = TRUE)

### rename source to be consistent with indirect
direct$contam <- ifelse(
  is.na(direct$source) == TRUE, "control",
                        ifelse(
                          direct$source ==
                            "effluents","RW",
                               "BS"))

direct$contam <- factor(direct$contam,
                        levels = c(
                          "control",
                          "RW",
                          "BS"
                        ))

### spiking level plus contam
direct$contam_spike <- paste0(direct$contam, 
                                "-", 
                                direct$spiking_level)

### treatment
direct$contam_spike <- factor(direct$contam_spike, 
                                levels = c("control-0",
                                           "RW-0",
                                           "RW-1",
                                           "RW-2",
                                           "BS-0",
                                           "BS-1",
                                           "BS-2"
                                           ),
                              ordered = TRUE)

## save formatted data file
save(direct, file = "./direct-formatted.Rda")
```

### Figures (raw data)

```{r figs_raw}
load(file = "./direct-formatted.Rda") ## loads direct

## plots (traits at harvest)

### traits to run through (all crops)
trait.list1 <- c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "shoot_root_ratio",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "germination_perc",
                "survival_perc"
                )

## source function
source("./Source_code/plot_raw_data_func.R")

## plotting function
plots.out1 <- sapply(trait.list1, 
                     plot_raw_data_func,
                     df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## peas only

### traits to run through (peas only)
trait.list2 <- c(
                "wet_pod_weight",
                "dry_pod_weight",
                "pods",
                "flowers",
                "infected_peas_perc"
                )

## source function
source("./Source_code/plot_raw_data.peas_func.R")

## plotting function
plots.out2 <- sapply(trait.list2,
                    plot_raw_data.peas_func,
                    df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## combine plots (all crops)
fig1 <- plot_grid(plots.out1[["dry_shoot_weight"]] +
                    labs(y = "Dry shoot \n weight (g)"),
                  plots.out1[["dry_root_weight"]] +
                    labs(y = "Dry root \n weight (g)"),
                  plots.out1[["shoot_root_ratio"]] +
                    labs(y = "Shoot:root \n ratio"),
                  plots.out1[["shoot_moisture"]] +
                    labs(y = "Shoot moisture"),
                  plots.out1[["survival_perc"]] +
                    labs(y = "Survival (%)"),
          ncol = 1,
          nrow = 5,
          align = "hv",
          labels = NULL)

## combine plots (peas only)
fig2 <- plot_grid(plots.out2[["dry_pod_weight"]] +
                    labs(y = "Dry pod \n weight (g)"),
                  plots.out2[["infected_peas_perc"]] +
                    labs(y = "Peas Infected (%)"),
          ncol = 2,
          nrow = 1,
          align = "hv",
          labels = NULL)

## combine plots
fig <- plot_grid(fig1, fig2,
          ncol = 1,
          nrow = 2,
          rel_heights = c(0.8,0.2),
          align = "hv",
          labels = NULL)

save_plot("./raw_data-direct.png", fig,
          ncol = 2, 
          nrow = 5,
          base_aspect_ratio = 1.3)
```

## Models

### Model 1

Impact of amendment type on plant traits, compared across 
different crop species

Include controls. 
Limit to Spiking level 0 (no additional contaminants added)

```{r model1}
## load data
load(file = "./direct-formatted.Rda") ## loads direct

## set contrasts for ANOVA
options(contrasts=c("contr.sum","contr.poly")) 

### add 1's to survival (to take log)
direct$survival_perc.corr <- direct$survival_perc + 1

### traits to run through (non-peas)
trait.list1 <- c(
                "wet_shoot_weight", 
                "dry_shoot_weight",
                "wet_root_weight", 
                "dry_root_weight",
                "shoot_root_ratio",
                "dry_total_weight",
                "per_plant_weight",
                "shoot_moisture",
                "germination_perc",
                "survival_perc.corr"
                )

## source function
source("./Source_code/lmm1_func.R")

## model
mod1.out <- sapply(trait.list1, 
                   lmm1_func,
                   df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## combine dfs
### ANOVAs
aov <- lapply(mod1.out, `[[`, 1) %>%
  bind_rows(.)
write.csv(aov, "./model_outputs/direct/aov1.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod1.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/direct/emm1.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod1.out, `[[`, 3) %>%
  bind_rows(.)
write.csv(cont, "./model_outputs/direct/cont1.csv", 
          row.names = FALSE)

## peas only

### add 1's to infected
direct$infected_peas_perc.corr <- 
  ifelse(is.na(direct$infected_peas_perc) == FALSE,
                                         direct$infected_peas_perc + 1,
                                         NA)

### traits to run through (peas only)
trait.list2 <- c("wet_pod_weight",
                "dry_pod_weight",
                "infected_peas_perc.corr",
                "pods",
                "flowers"
                )

## source function
source("./Source_code/lmm1.peas_func.R")

## model
mod1_peas.out <- sapply(trait.list2, 
                   lmm1.peas_func,
                   df = direct,
                   simplify = FALSE, 
                   USE.NAMES = TRUE)

## combine dfs
### ANOVAs
aov <- lapply(mod1_peas.out, `[[`, 1) %>%
  bind_rows(.)
write.csv(aov, "./model_outputs/direct/aov_peas.csv", 
          row.names = FALSE)
### emmeans
emm <- lapply(mod1_peas.out, `[[`, 2) %>%
  bind_rows(.)
write.csv(emm, "./model_outputs/direct/emm_peas.csv", 
          row.names = FALSE)
### CONTRASTS
cont <- lapply(mod1_peas.out, `[[`, 3) %>%
  bind_rows(.)
write.csv(cont, "./model_outputs/direct/cont_peas.csv", 
          row.names = FALSE)

## figures (cont)
cont.all <- read_csv("model_outputs/direct/cont1.csv")
cont.peas <- read_csv("model_outputs/direct/cont_peas.csv")
cont.peas$species <- "pea"
cont.peas$lower.CL <- ifelse(is.na(cont.peas$lower.CL) == TRUE,
                             cont.peas$asymp.LCL,
                             cont.peas$lower.CL)
cont.peas$upper.CL <- ifelse(is.na(cont.peas$upper.CL) == TRUE,
                             cont.peas$asymp.UCL,
                             cont.peas$upper.CL)
cont.peas$t.ratio <- ifelse(is.na(cont.peas$t.ratio) == TRUE,
                             cont.peas$z.ratio,
                             cont.peas$t.ratio)
cont.peas <- cont.peas %>%
  select(-asymp.LCL, -asymp.UCL, -z.ratio)

cont <- rbind(cont.all, cont.peas)

cont$contam <- ifelse(grepl("BS", cont$contrast, fixed = FALSE),
                      "BS", "RW")

cont$contam <- factor(cont$contam,
                      levels = c("RW",
                                 "BS"))

## subset to sig traits
cont.f <- cont %>%
  filter(trait %in% c("dry_shoot_weight","dry_root_weight",
         "shoot_root_ratio","shoot_moisture",
         "germination_perc","survival_perc.corr","pods")) %>%
           droplevels(.)

cont.f$trait <- factor(cont.f$trait,
          levels = c("dry_shoot_weight","dry_root_weight",
         "shoot_root_ratio","shoot_moisture",
         "germination_perc","survival_perc.corr","pods"))

trait_names <- c(
  dry_shoot_weight = 'Dry shoot \n weight (g)',
  dry_root_weight = 'Dry root \n weight (g)',
  shoot_root_ratio = 'Shoot to root \n ratio',
  shoot_moisture = 'Shoot moisture \n (%)',
  germination_perc = 'Germination \n (%)',
  survival_perc.corr = 'Survival \n (%)',
  pods = "Pods (no.)"
)

species_names <- c(
  lettuce = 'Lettuce',
  pea = 'Pea',
  radish = 'Radish'
)

### plot
ggplot(data = cont.f,
   aes(x = contam, 
       y = ratio, 
       colour = contam)) +
  geom_pointrange(aes(ymin = lower.CL,
                      ymax = upper.CL),
                  position = position_dodge(0.5)) +
  geom_hline(aes(yintercept = 1),
             linetype = 2) +
   scale_colour_manual(values = c(
                                  "blue",
                                 "#B06500")) +
  facet_grid(trait~species, scales = "free",
             labeller = labeller(trait = trait_names,
                                 species = species_names)) +
  labs(y = NULL, x = NULL) +
  guides(colour = "none") +
  theme_bw() +
  theme(
    axis.text.x = element_text(),
    strip.text = element_text(face = "bold"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )

ggsave("./model_outputs/direct/cont1.png", width = 6, height = 12, units = "in")

```
